---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Lab6

<!-- badges: start -->
<!-- badges: end -->

The goal of Lab6 is to ...

## Installation

You can install the development version of Lab6 from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("seymi176/Lab6")
```

## 1. Brute force method

Implementation of brute_force method for solving the knapsack problem :

```{r , eval = T}
library(Lab6)
data(knapsack_objects)
brute_force_knapsack(x = knapsack_objects[1:8,], W = 3500)
brute_force_knapsack(x = knapsack_objects[1:12,], W = 3500)
brute_force_knapsack(x = knapsack_objects[1:8,], W = 2000)
brute_force_knapsack(x = knapsack_objects[1:12,], W = 2000)
```

## 2. Dynamic programming
Implementation of dynamic programming method for solving the knapsack problem :

```{r , eval = T}
data(knapsack_objects)
knapsack_dynamic(x = knapsack_objects[1:8,], W = 3500)
knapsack_dynamic(x = knapsack_objects[1:12,], W = 3500)
knapsack_dynamic(x = knapsack_objects[1:8,], W = 2000)
knapsack_dynamic(x = knapsack_objects[1:12,], W = 2000)
```

## 3. Greedy heuristic
Implementation of Greedy heuristic method for solving the knapsack problem :

```{r , eval = T}
data(knapsack_objects)
greedy_knapsack(x = knapsack_objects[1:800,], W = 3500)
greedy_knapsack(x = knapsack_objects[1:1200,], W = 2000)
```