---
title: "Writing fast R code"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Writing fast R code}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
body {
text-align: justify}
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This package is created to study the effects of algorithms with different computational
complexity and investigate how to speedup R code.The package aims to solve the NP-hard knapsack problem by defining three functions, each utilizing different approach and method. The knapsack problem is a discrete optimization problem where we have a knapsack that can take a
limited weight $W$ and we want to fill this knapsack with a number of items *i* = 1, ..., n, each with a weight w~i~ and a value v~i~. The goal is to find the knapsack with the largest value of the elements added to the knapsack.


### 1.1.2 Brute force search

The brute_force_knapsack function searches through all possible options and returns the highest value discovered. It provides the right response, but its complexity is O(2^n) because all possible (2^n) possibilities must be considered.

As shown below, it takes less than *10* seconds for the function *"brute_force_knapsack"* to find the solution for n = 22 objects.

```{r setup}
library(Lab6)
system.time(brute_force_knapsack(x = knapsack_objects[1:22,], W = 2000), gcFirst = T)
```


### 1.1.3 Dynamic programming
The knapsack_dynamic is based on the fact that all weights are nonnegative integers and they are all less than $W$. In this approach it's defined a matrix where in each position m[i,w] is stored the maximum value that can be attained with weight less than or equal to w using items up to i. The maximum value is found calculating m[n,W]. It gives a correct answer to the problem in all the situation but its complexity is $O(Wn)$ less than the brute_force_knapsackfunction.

As expected, the dynamic algorithm has better performance than previous one, and it takes around *10* seconds for the  *"knapsack_dynamic"* to find the solution for n = 500 objects.

```{r}
system.time(knapsack_dynamic(x = knapsack_objects[1:500,], W = 3500), gcFirst = T)
```

### 1.1.4 Greedy heuristic

In terms of calculation time, the greedy algorithm has the best performance among others. It takes around *1* second for the  *"greedy_knapsack"* to find the solution for n = 1000000 objects. First, we create a new knapsack object which contain 1000000 objects and then test performance of the algorithm.

```{r}
RNGversion(min(as.character(getRversion()),"3.5.3"))
set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")
n2 <- 1000000
knapsack_objects_new <-
data.frame(
w=sample(1:4000, size = n2, replace = TRUE),
v=runif(n = n2, 0, 10000)
)

system.time(greedy_knapsack(x = knapsack_objects_new[1:1000000,], W = 3500), gcFirst = T)
```

### 1.1.5 test suits

test suits for the knapsack_brute_force function:

```{r , eval=FALSE}

suppressWarnings(RNGversion(min(as.character(getRversion()),"3.5.3")))
set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")
n <- 2000
knapsack_objects <- data.frame(w=sample(1:4000, size = n, replace = TRUE), v=runif(n = n, 0, 10000))


test_that("Correct object is returned", {
  expect_silent(bfk <- brute_force_knapsack(x = knapsack_objects[1:8,], W = 3500))
  expect_named(bfk, c("value", "elements"))
})


test_that("functions rejects errounous input.", {
  expect_error(brute_force_knapsack("hej", 3500))
  expect_error(brute_force_knapsack(x = knapsack_objects[1:8,], W = -3500))
})

test_that("Function return correct results.", {
  bfk <- brute_force_knapsack(x = knapsack_objects[1:8,], W = 3500)
  expect_equal(round(bfk$value), 16770)
  expect_true(all(round(bfk$elements) %in% c(5, 8)))

  bfk <- brute_force_knapsack(x = knapsack_objects[1:12,], W = 3500)
  expect_equal(round(bfk$value), 16770)
  expect_true(all(round(bfk$elements) %in% c(5, 8)))

  bfk <- brute_force_knapsack(x = knapsack_objects[1:8,], W = 2000)
  expect_equal(round(bfk$value), 15428)
  expect_true(all(round(bfk$elements) %in% c(3, 8)))

  bfk <- brute_force_knapsack(x = knapsack_objects[1:12,], W = 2000)
  expect_equal(round(bfk$value), 15428)
  expect_true(all(round(bfk$elements) %in% c(3, 8)))

  st <- system.time(bfk <- brute_force_knapsack(x = knapsack_objects[1:16,], W = 2000))
  expect_true(as.numeric(st)[2] >= 0.00)
})
```

test suits for the greedy_knapsack function:
```{r , eval=FALSE}
suppressWarnings(RNGversion(min(as.character(getRversion()),"3.5.3")))
set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")
n <- 2000
knapsack_objects <- data.frame(w=sample(1:4000, size = n, replace = TRUE), v=runif(n = n, 0, 10000))

test_that("Correct object is returned", {
  expect_silent(gk <- greedy_knapsack(x = knapsack_objects[1:8,], W = 3500))
  expect_named(gk, c("value", "elements"))
})

test_that("functions rejects errounous input.", {
  expect_error(greedy_knapsack("hej", 3500))
  expect_error(greedy_knapsack(x = knapsack_objects[1:8,], W = -3500))
})

test_that("Function return correct results.", {
  gk <- greedy_knapsack(x = knapsack_objects[1:8,], W = 3500)
  expect_equal(round(gk$value), 15428)
  expect_true(all(round(gk$elements) %in% c(3, 8)))
  
  gk <- greedy_knapsack(x = knapsack_objects[1:12,], W = 3500)
  expect_equal(round(gk$value), 15428)
  expect_true(all(round(gk$elements) %in% c(3, 8)))
  
  gk <- greedy_knapsack(x = knapsack_objects[1:8,], W = 2000)
  expect_equal(round(gk$value), 15428)
  expect_true(all(round(gk$elements) %in% c(3, 8)))
  
  gk <- greedy_knapsack(x = knapsack_objects[1:12,], W = 2000)
  expect_equal(round(gk$value), 15428)
  expect_true(all(round(gk$elements) %in% c(3, 8)))
  
  st <- system.time(gk <- greedy_knapsack(x = knapsack_objects[1:16,], W = 2000))
  expect_true(as.numeric(st)[2] <= 0.01)
  
  gk <- greedy_knapsack(x = knapsack_objects[1:800,], W = 3500)
  expect_equal(round(gk$value), 192647)
  
  gk <- greedy_knapsack(x = knapsack_objects[1:1200,], W = 3500)
  expect_equal(round(gk$value), 270290)
})
```

test suits for the knapsack_dynamic function:

```{r , eval=FALSE}
suppressWarnings(RNGversion(min(as.character(getRversion()),"3.5.3")))
set.seed(42, kind = "Mersenne-Twister", normal.kind = "Inversion")
n <- 2000
knapsack_objects <- data.frame(w=sample(1:4000, size = n, replace = TRUE), v=runif(n = n, 0, 10000))

test_that("Correct object is returned", {
  expect_silent(dk <- knapsack_dynamic(x = knapsack_objects[1:8,], W = 3500))
  expect_named(dk, c("value", "elements"))
})

test_that("functions rejects errounous input.", {
  expect_error(knapsack_dynamic("hej", 3500))
  expect_error(knapsack_dynamic(x = knapsack_objects[1:8,], W = -3500))
})

test_that("Function return correct results.", {
  dk <- knapsack_dynamic(x = knapsack_objects[1:8,], W = 3500)
  expect_equal(round(dk$value), 16770)
  expect_true(all(dk$elements %in% c(5, 8)))

  dk <- knapsack_dynamic(x = knapsack_objects[1:12,], W = 3500)
  expect_equal(round(dk$value), 16770)
  expect_true(all(dk$elements %in% c(5, 8)))

  dk <- knapsack_dynamic(x = knapsack_objects[1:8,], W = 2000)
  expect_equal(round(dk$value), 15428)
  expect_true(all(dk$elements %in% c(3, 8)))

  dk <- knapsack_dynamic(x = knapsack_objects[1:12,], W = 2000)
  expect_equal(round(dk$value), 15428)
  expect_true(all(dk$elements %in% c(3, 8)))
   
  st <- system.time(dk <- knapsack_dynamic(x = knapsack_objects[1:16,], W = 2000))
  expect_true(as.numeric(st)[2] <= 0.01)

  dk <- knapsack_dynamic(x = knapsack_objects[1:24,], W = 3500)
  expect_equal(round(dk$value), 28828) #ERROR
})
```


### 1.1.8 Parallelize brute force search

As shown below, it takes less time for the function, compared to the non-parallelized method.

```{r }
library(Lab6)
system.time(brute_force_knapsack(x = knapsack_objects[1:22,], W = 2000, parallel=TRUE),gcFirst = T)
```

